\chapter{Recruitment note}\label{app:recruitment-note}

The following note was shared by us in collaboration with out supervisor and co-supervisor on the NTNU intranet, ``Innsida'', and with members of EXCITED. The purpose was to recruit instructors to help us collect open-text responses from their students.

\setlength\parskip{1em plus 0.1em minus 0.2em}
\setlength\parindent{0pt}

\section*{Looking for teachers to support the ``AI in the classroom'' master's thesis}

\textbf{We are looking for teachers to support the "AI in the classroom" master's thesis. In return we are hoping to help you to support you in your lectures. If you are already using interactive tools in your lectures, you could help us by sharing the anonymized open text answers from your students. More details and contact information is below.}

\textbf{Are you looking to boost student engagement and deepen learning outcomes in your courses?}

We are two students working on our master's thesis focused on improving in-classroom interactivity using open-ended questions. We are seeking lecturers interested in exploring this approach and collaborating with us on our research.

\textbf{Purpose of Our Master Thesis}

\begin{itemize}
    \item \textbf{Enhancing Interactivity and Reflection:} Open-ended questions encourage students to reflect more deeply, leading to more thoughtful responses and honest feedback.
    \item \textbf{Real-Time Challenges:} We recognize that interpreting numerous open-text responses during a lecture can be challenging.
    \item \textbf{Our Goal:} To analyze and experiment with these open-ended responses to develop effective strategies for real-time interpretation and feedback.
\end{itemize}

\textbf{Value for You as a Lecturer}

\begin{itemize}
    \item \textbf{Increase Engagement:} Implementing open-ended questions can make your lectures more interactive.
    \item \textbf{Early Misconception Detection:} Identify and address student misunderstandings promptly.
    \item \textbf{Insightful Feedback:} Open-ended responses may provide you with valuable feedback to improve your course.
\end{itemize}

\textbf{What Participation Involves}

\begin{itemize}
    \item \textbf{Incorporate Questions:} Integrate a few open-ended questions into your lectures that require 2-5 minutes for students to answer.
    \item \textbf{Collect Responses:} Use tools like Mentimeter or Kahoot to gather anonymous student responses.
    \item \textbf{Use the responses:} Optionally use the insight to adapt or improve your lectures in real-time.
    \item \textbf{Share Data:} Email the collected responses to us.
\end{itemize}

\textbf{Next Steps}

If this collaboration interests you, please contact us. We would be delighted to arrange a meeting at your convenience to discuss details — without any obligation on your part.

Your participation would not only contribute significantly to our research but also potentially enhance your teaching practice.

Best regards,

Sander Østrem Fagernes (sanderof@stud.ntnu.no)

Fredrik Fonn Hansen (fredrik.f.hansen@ntnu.no)

\setlength\parskip{0pt plus 1pt}
\setlength\parindent{15pt}